---
title: "Phylopars visualization"
author: "April Lamb"
date: "7/19/2021"
output: html_document
---

```{r, include = FALSE}
# Function to reduce line output of markdown file
# save the built-in output hook
hook_output <- knitr::knit_hooks$get("output")

# set a new output hook to truncate text output
knitr::knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$out.lines)) {
    x <- xfun::split_lines(x)
    if (length(x) > n) {
      # truncate the output
      x <- c(head(x, n), "....\n")
    }
    x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})
```

### Load libraries and read in data ###
```{r setup, include=FALSE, out.lines = 4}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(ggbeeswarm)
library(data.table)

# Set working directory. File path will need to be updated
setwd('C:/Users/april/OneDrive/Desktop/Covid/phylopars_updated_7-22-21') 
data <- read.csv("CoV2_Imputations.csv")
data$month <- as.factor(data$month)
data$nonrecombinant <- as.factor(data$nonrecombinant)
```

### Parse data by site & alignment type ###
```{r split 1, out.lines = 4}
# Isolate each location
(Beijing    <- data[data$site == "Beijing", ])
(Tampere    <- data[data$site == "Tampere", ])
(Guangzhou  <- data[data$site == "Guangzhou", ])
(HongKong   <- data[data$site == "Hong Kong", ])
(Sarlahi    <- data[data$site == "Sarlahi", ])
(Amsterdam  <- data[data$site == "Amsterdam", ])
(Norway     <- data[data$site == "Sor Trondelag", ])
(NY         <- data[data$site == "New York", ])
(Japan      <- data[data$site == "Yamagata", ])
(Colorado   <- data[data$site == "Colorado", ])
(Gothenburg <- data[data$site == "Gothenburg", ])
(Stockholm  <- data[data$site == "Stockholm", ])
(Thailand   <- data[data$site == "Nakhon Si Thammarat", ])
(SKorea     <- data[data$site == "South Korea", ])

# Group NA + Europe & Asia locales together
ENA  <- rbind(NY, Colorado, Stockholm, Gothenburg, Amsterdam, Tampere, Norway)
Asia <- rbind(Beijing, Guangzhou, HongKong, Sarlahi, Japan, Thailand, SKorea)

# split further by alignment type
(ENAnonrecom   <- ENA[ENA$nonrecombinant == "yes", ])
(ENArecom      <- ENA[ENA$nonrecombinant == "no", ])
(Asianonrecom  <- Asia[Asia$nonrecombinant == "yes", ])
(Asiarecom     <- Asia[Asia$nonrecombinant == "no", ])

# Pull out the monthly average for each site for plotting bar charts
ENA_nonaverage  <- aggregate(ENAnonrecom$cases, list(ENAnonrecom$month, ENAnonrecom$site), mean)
ENA_average     <- aggregate(ENArecom$cases, list(ENArecom$month, ENArecom$site), mean)
Asia_nonaverage <- aggregate(Asianonrecom$cases, list(Asianonrecom$month, Asianonrecom$site), mean)
Asia_average    <- aggregate(Asiarecom$cases, list(Asiarecom$month, Asiarecom$site), mean)

# Rename columns
ENA_nonaverage <- setnames(ENA_nonaverage, old = c('Group.1','Group.2', 'x'), new = c('month','site', 'avg.cases'))
ENA_average <- setnames(ENA_average, old = c('Group.1','Group.2', 'x'), new = c('month','site', 'avg.cases'))
Asia_nonaverage <- setnames(Asia_nonaverage, old = c('Group.1','Group.2', 'x'), new = c('month','site', 'avg.cases'))
Asia_average <- setnames(Asia_average, old = c('Group.1','Group.2', 'x'), new = c('month','site', 'avg.cases'))

```

### Plot monthly SARS-CoV-2 proportions by site & facet by tree type ###

Beijing
```{r Beijing}
#pdf("Beijing_CoV.pdf")
ggplot() + geom_col(data = Beijing, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Tampere, Finland
```{r Tampere}
#pdf("Tampere_CoV.pdf")
ggplot() + geom_col(data = Tampere, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Guangzhou, China
```{r Guangzhou}
#pdf("Guangzhou_CoV.pdf")
ggplot() + geom_col(data = Guangzhou, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Hong Kong
```{r HongKong}
#pdf("HongKong_CoV.pdf")
ggplot() + geom_col(data = HongKong, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Sarlahi, Nepal
```{r Sarlahi}
#pdf("Sarlahi_CoV.pdf")
ggplot() + geom_col(data = Sarlahi, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, Sarlahi", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Amsterdam
```{r Amsterdam}
#pdf("Amsterdam_CoV.pdf")
ggplot() + geom_col(data = Amsterdam, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, Amsterdam", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Trondelag, Norway
```{r Norway}
#pdf("Norway_CoV.pdf")
ggplot() + geom_col(data = Norway, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, Norway", x = "Month", y = "Percent") + 
  facet_wrap(~ tree) 
#dev.off()
```

Thailand
```{r Thailand}
#pdf("Thailand_CoV.pdf")
ggplot() + geom_col(data = Thailand, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, Thailand", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Colorado
```{r Co}
#pdf("Colorado_CoV.pdf")
ggplot() + geom_col(data = Colorado, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, Colorado", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

New York
```{r NY}
#pdf("NY_CoV.pdf")
ggplot() + geom_col(data = NY, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, New York", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Japan
```{r Japan}
#pdf("Japan_CoV.pdf")
ggplot() + geom_col(data = Japan, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, Yamagata", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Gothenburg, Sweden
```{r Sweden1}
#pdf("Gothenburg_CoV.pdf")
ggplot() + geom_col(data = Gothenburg, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, Gothenburg", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Stockholm, Sweden
```{r Sweden2}
#pdf("Stockholm_CoV.pdf")
ggplot() + geom_col(data = Stockholm, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, Stockholm", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

Nationwide, South Korea
```{r SKorea}
#pdf("SKorea_CoV.pdf")
ggplot() + geom_col(data = SKorea, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Proportion of Sars-Cov-2 cases by tree, South Korea", x = "Month", y = "Percent") + 
  facet_wrap(~ tree)
#dev.off()
```

### Check sensitivity to method of time tree inference and use of only viral sequences deemed non-recombining ###

Europe+North America, nonrecombinant
```{r ENA nonrecomb}
# Plot points
#pdf("ENA_nonrecomb_beeswarm.pdf")
ggplot(ENAnonrecom, aes(month, cases, color=tree)) +
    geom_beeswarm(cex=1, size=.25) +
    facet_wrap(site~.) +
    ggtitle("ENA-nonrecom") +
    xlab("Month") +
    ylab("Parameter estimate") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
#dev.off()

# Plot averages
#pdf("ENA_nonaverage.pdf")
ggplot(ENA_nonaverage, aes(month, avg.cases)) + 
    geom_bar(stat="identity") +
    facet_wrap(site~.) + 
    ylim(0,25)
#dev.off()
```

Europe+North America, recombinant
```{r ENA recomb}
# Plot points 
#pdf("ENA_recomb_beeswarm.pdf")
ggplot(ENArecom, aes(month, cases, color=tree)) +
    geom_beeswarm(cex=1, size=.25) +
    facet_wrap(site~.) +
    ggtitle("ENA-recom") +
    xlab("Month") +
    ylab("Parameter estimate") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
#dev.off()

# Plot averages
#pdf("ENA_average.pdf")
ggplot(ENA_average, aes(month, avg.cases)) + 
    geom_bar(stat="identity") +
    facet_wrap(site~.) + 
    ylim(0,25)
#dev.off()
```

Asia, nonrecombinant
```{r Asia nonrecomb}
# Plot points
#pdf("Asia_nonrecomb_beeswarm.pdf")
ggplot(Asianonrecom, aes(month, cases, color=tree)) +
    geom_beeswarm(cex=1, size=.25) +
    facet_wrap(site~.) +
    ggtitle("Asia-nonrecom") +
    xlab("Month") +
    ylab("Parameter estimate") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
#dev.off()

# Plot averages
#pdf("Asia_nonaverageTest.pdf")
ggplot(Asia_nonaverage, aes(month, avg.cases)) + 
    geom_bar(stat="identity") +
    facet_wrap(site~.) + 
    ylim(0,40)
#dev.off()
```

Asia, recombinant
```{r Asia recomb}
# Plot points
#pdf("Asia_recomb_beeswarm.pdf")
ggplot(Asiarecom, aes(month, cases, color=tree)) +
    geom_beeswarm(cex=1, size=.25) +
    facet_wrap(site~.) +
    ggtitle("Asia-recom") +
    xlab("Month") +
    ylab("Parameter estimate") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
#dev.off()

# Plot averages
#pdf("Asia_average.pdf")
ggplot(Asia_average, aes(month, avg.cases)) + 
    geom_bar(stat="identity") +
    facet_wrap(site~.) + 
    ylim(0,30)
#dev.off()
```

## Now let's plot the endemic coronaviruses at each site
```{r data, out.lines = 4}
setwd('C:/Users/april/OneDrive/Desktop/Covid/phylopars_updated_7-22-21') # File path will need to be updated
df <- read.csv("CoV2_Imputations_IQTREE_time_tree.csv", stringsAsFactors = TRUE) # read data
df$month <- as.factor(df$month)

(cov229E  <- df[df$type == "229E", ])
(covOC43  <- df[df$type == "OC43", ])
(covNL63  <- df[df$type == "NL63", ])
(covHKU1  <- df[df$type == "HKU1", ])
```

229E
```{r 229E}
#pdf("229E_all.pdf")
ggplot() + geom_col(data = cov229E, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Relative monthly cases by coronavirus- 229E", x = "Month", y = "Percent") + 
  facet_wrap(~ site)+
  ylim(0,40)
#dev.off()
```

OC43
```{r OC43}
#pdf("OC43_all.pdf")
ggplot() + geom_col(data = covOC43, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Relative monthly cases by coronavirus- OC43", x = "Month", y = "Percent") + 
  facet_wrap(~ site)+
  ylim(0,40)
#dev.off()
```

NL63
```{r NL63}
#pdf("NL63_all.pdf")
ggplot() + geom_col(data = covNL63, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Relative monthly cases by coronavirus- NL63", x = "Month", y = "Percent") + 
  facet_wrap(~ site)+
  ylim(0,40)
#dev.off()
```

HKU1
```{r HKU1}
#pdf("HKu1_all.pdf")
ggplot() + geom_col(data = covHKU1, aes(x = month, y = cases), position = "dodge") + 
  labs(title= "Relative monthly cases by coronavirus- HKU1", x = "Month", y = "Percent") + 
  facet_wrap(~ site)+
  ylim(0,40)
#dev.off()
```




























